Tropopause Evolution in a Rapidly Intensifying Tropical Cyclone: A Static Stability Budget Analysis in an Idealized, Axisymmetric Framework

### 2018-02-27 paper.tex
Added Appendix figures and expanded a lot of the Discussion and some of the Appendix.

\abstract{Large changes in tropopause-layer static stability are observed during the rapid intensification (RI) of an idealized, axisymmetric tropical cyclone (TC).
Over the eye, static stability near the tropopause decreases and the cold-point tropopause height rises by up to 4 km at the storm center.
Outside of the eye, static stability immediately above the cold-point tropopause increases considerably, and the tropopause remains near its initial level.
A budget analysis reveals that advection contributes to the static stability tendencies at all times throughout the upper troposphere and lower stratosphere.
Advection is particularly important within the eye, where it acts to destabilize the layer near and above the cold-point tropopause.
Outside of the eye, a radial-vertical circulation develops during RI, with strong outflow below the tropopause and weak inflow above.
Vertical wind shear above and below the upper-tropospheric outflow maximum induces turbulence, which provides forcing for both destabilization and stabilization in the tropopause layer.
Meanwhile, as organized convection reaches the tropopause, radiative heating tendencies at the top of the cirrus canopy generally act to destabilize the upper troposphere and stabilize the lower stratosphere.
Turbulent mixing and radiative heating combine to play an important role in the development of the strong stable layer immediately above the cold-point tropopause during RI.
The results suggest that turbulence and radiation, alongside advection, play fundamental roles in the upper-level static stability evolution of TCs.}
%The conclusions are robust to changes in the initial thermodynamic profile and prescribed vertical mixing length within a reasonable range of values.}

@@

Idealized simulations of a TC analyzed by \cite{OhnoSatoh2015} suggested that the development of an upper-level warm core near the TC storm center acted to decrease the static stability near the tropopause (see their Fig. ).
Although the mechanisms that drive this static stability evolution have not been examined explicitly, \cite{SternZhang2013} described the development of the TC warm core using a potential temperature ($\theta$) budget analysis.
They found that radial and vertical advection both played important roles in warm core development throughout RI, and subgrid-scale diffusion became particularly important during the later stage of RI.
To our knowledge, the only paper that has examined explicitly the static stability evolution in a modeled TC is \cite{Kepertetal2016}, but their analysis was limited to the boundary layer.
The analysis herein is based upon that of \cite{SternZhang2013}, except using a static stability budget similar to that of \cite{Kepertetal2016}, with a focus on the upper troposphere and lower stratosphere.

@@

In the tropopause layer, some of the budget terms are small enough to be ignored.
To determine which of the budget terms are most important, a time series of the contribution of each of the budget terms in Eq. \ref{eq:dthetadt} to the tropopause-layer static stability tendency is plotted in Fig. \ref{fig:avgbudterms}.
For this figure, each of the budget terms is computed using the method described in Section 3, except with 1-hour averaging intervals instead of 24-hour intervals.
The absolute values of these tendencies are then averaged over a radius-height domain surrounding the tropopause and plotted as a time series\footnote{It will be seeen in subsequent figures that each of the terms contributes both positively and negatively to the $N^2$ tendency within the analysis domain. 
Thus, taking an average over the domain tends to wash out the positive and negative contributions.
To circumvent this problem, the absolute value of each of the terms is averaged.}. 
Advection (Fig.~\ref{fig:avgbudterms}, red line) plays an important role in the mean tropopause-layer static stability tendency at all times, and vertical turbulence (Fig.~\ref{fig:avgbudterms}, blue line) and radiation (Fig.~\ref{fig:avgbudterms}, dark green line) also contribute significantly. %both become important after 48 hours.
Although the contribution from horizontal turbulence (Fig.~\ref{fig:avgbudterms}, purple line) becomes more important after 48 hours, it is confined to a very small region immediately surrounding the eyewall tangential velocity maximum (not shown), and is negligible throughout the rest of the tropopause layer.
The remaining two processes - microphysics and dissipative heating (Fig.~\ref{fig:avgbudterms}, orange and light green lines, respectively) - lie atop one another near zero.
These time series indicate that, at all times, three budget terms dominate the tropopause-layer static stability tendency: advection, vertical turbulence, and radiation.
Variations in the magnitude and spatial structure of these terms drive the static stability changes depicted in Fig.~\ref{fig:mod+bud+res}; subsequent sections will focus on these variations and what causes them.

@@

\paragraph{0-24 hours}
The weakening of the lower-stratospheric static stability maximum during the initial spin-up period is reflected in the total $N^2$ budget change over this time (Fig.~\ref{fig:stab-00-24}a).
The 17-18-km layer was characterized by decreasing $N^2$ (purple shading), maximizing at the storm center.
The layer immediately below the tropopause, meanwhile, saw strengthening $N^2$ during this time period.
Although these tendencies extended out to the 200-km radius, they were particularly pronounced at innermost radii.
A comparison of the contributions of advection (Fig.~\ref{fig:stab-00-24}b), vertical turbulence (Fig.~\ref{fig:stab-00-24}c), and radiation (Fig.~\ref{fig:stab-00-24}d) reveals that advection is primarily responsible for the change in static stability during this period.
Although vertical turbulence acts in opposition to advection (i.e. it acts to stabilize regions that advection acts to destabilize), the magnitude of the advective tendencies is larger, particularly at the innermost radii.
The sum of advection and vertical turbulence (Fig.~\ref{fig:stab-00-24}e) almost exactly replicates the static stability tendencies above 17 km.
Radiative tendencies (Fig.~\ref{fig:stab-00-24}d) act to destabilize the layer below about 16 km and stabilize the layer between 16 and 17 km.
The sum of advection, vertical turbulence, and radiation (Fig.~\ref{fig:stab-00-24}f) reproduces the total change in $N^2$ almost exactly.

...Explain this in the context of radial and vertical velocities...
...See Stern and Zhang, Page 84, Section 3d...
...Add mention of total condensate and radiative heating tendencies as it relates to statbility tendency due to rad...

@@

TWO REGIONS WHERE Panel (f) differs from panel (a): 30-60 km radial band below 16 km, which is actually canceled out by a vertical gradient of latent heating, and the thin region of strong stabilization between 15-17.5 km near r=30 km, which is canceled out by horizontal turbulence.

@@

Advection played an important role in the tropopause-layer $N^2$ evolution at all stages of intensification, but for brevity, this section will focus only on the RI (24-48-hour) period.
To investigate the advective processes more closely, the individual contributions of horizontal and vertical advection during the RI period are shown in Fig. \ref{fig:adv-24-48}, along with the corresponding time-mean radial and vertical velocities and $\theta$.
The $N^2$ tendencies due to the two advective components (Fig. \ref{fig:adv-24-48}a,b) exhibit strong cancellation, consistent with flow that is nearly isentropic.
There are, however, some regions in which flow crosses $\theta$ surfaces; this flow accounts for all non-zero $N^2$ tendencies due to advection previously seen in Fig. \ref{fig:stab-24-48}b.

@@

The first two terms on the right-hand side of Eq. \ref{eq:advtend} represent advection of static stability by the radial and vertical wind, respectively.
These terms cannot create a maximum or a minimum; they only act to rearrange the static stability field.
Since the goal of this analysis is to understand the processes responsible for the creation of maxima and minima in $N^2$, these two terms will not be examined.
The third and fourth terms represent, respectively, the tilting of isentropes in the presence of vertical wind shear, and the spreading or compaction of isentropes through divergence of the vertical wind.
Since these terms involve gradients of velocities, they can create or eliminate local maxima or minima in static stability. This section will focus on these two terms.

@@

%   \paragraph{RI period: 24-48 hours}
During the RI period, strong radial and vertical circulations developed near the tropopause, which forced high-magnitude $N^2$ tendencies due to advection (Fig. \ref{fig:adv-24-48}a,b).
A layer of strong outflow developed at and below the tropopause during this period, with the outflow maximum (dashed cyan line) curving from the 14-km level at the 50-km radius to just below the 16-km level outside of the 80-km radius (Fig. \ref{fig:adv-24-48}c).
The cyan line, by definition, represents the level at which the vertial gradient of radial velocity switched signs, with $\partial u/\partial z>0$ below the line and $\partial u/\partial z<0$ above.
Notably, the $N^2$ tendency due to horizontal advection (Fig. \ref{fig:adv-24-48}a) also tended to switch signs very near this line, with stabilization below the outflow maximum and destabilization above.
This suggests that vertical wind shear above and below the outflow maximum played an important role in the $N^2$ tendency during this time.
Examination of the third term on the right-hand side of Eq. \ref{eq:advtend} reveals that tilting of isentropes within these shear layers contributed to the strong $N^2$ tendencies that flanked the outflow maximum.
Outside of the eye and eyewall, isentropes generally sloped upward with radius, which means that $\theta$ decreased outward ($\partial \theta/\partial r<0$).
Thus, wherever $\partial u/\partial z>0$, the tilting term must force an increase in $N^2$, and wherever $\partial u/\partial z<0$, the tilting term must force a decrease in $N^2$.
This is precisely the structure seen in Fig. \ref{fig:adv-24-48}a, which suggests that the tilting term might be important near the outflow jet.

Advection of $N^2$ by the radial wind (first term on the right-hand side of Eq. \ref{eq:advtend}) also acts within the outflow jet.
For example, horizontal advection provided forcing for destabilization at the 16-km level almost everywhere inside of the 140-km radius.
Outside of this radius near 16 km, however, existed a region of forcing for stabilization.
This switch in signs was a consequence of a reversal of the radial graident of mean $N^2$ near the 140-km radius (Fig. \ref{fig:n2-24hr-avgs}b).
Inside of that radius, $(\partial/\partial r)(\partial \theta/\partial z) > 0$ and $u > 0$, which corresponds to forcing for destabilization in Eq. \ref{eq:advtend}; outside of that radius $(\partial/\partial r)(\partial \theta/\partial z) < 0$ and $u > 0$, which corresponds to forcing for stabilization.

@@

Meanwhile in the lower stratosphere, a thin layer of 2-4 m s\textsuperscript{-1} inflow developed a few hundred meters above the tropopause, similar to that which was observed in Hurricane Patricia (2015; \citeauthor{DuranMolinari2018} \citeyear{DuranMolinari2018}).
Since the isentropes in this layer slope slightly upward with radius (i.e. $\partial \theta/\partial r < 0$), this inflow acts to import lower $\theta$ air to inner radii.
There also existed strong radial gradients of mean $N^2$ in this layer during this time (Fig. \ref{fig:n2-24hr-avgs}b), so the inflow layer must have acted to rearrange $N^2$.
This rearrangement could not have been responsible for the overall increase in $N^2$ during this period, however.
If horizontal advection contributed to the strengthening stable layer just above the tropopause, it must have been through the tilting term.
The inflow maximum near the 18-km level represents the level at which the sign of $\partial u/\partial z$ reverses, with $\partial u/\partial z < 0$ below and $\partial u/\partial z > 0$ above.
The third term on the right-hand side of Eq. \ref{eq:advtend} suggests that the tilting of isentropes should act to destabilize the layer below the inflow maximum and stabilize the layer above, and this was observed in Fig. \ref{fig:adv-24-48}a.

@@

After the TC's RI period completed (48-72 hours; Fig. \ref{fig:qi+radten}f), strong radiative cooling remained near the tropopause at inner radii, sloping downward with the top of the cirrus canopy to below the tropopause at outer radii.
Cooling rates exceeded 1 K h\textsuperscript{-1} just above the tropopause between the 30- and 70-km radii.
These cooling rates exceeded those observed by \cite{buetal}, a discrepancy that is likely a consequence of their larger vertical grid spacing (625 m) compared to that used here (250 m), with a possible a contribution from differing radiation schemes\footnote{\cite{buetal} employed the NASA-Goddard radiation scheme for their CM1 simulations, whereas RRTMG is used in the present paper. They did use the RRTMG radiation scheme alongside the Thompson microphysics parameterization for their HWRF simulations, but those simulations also produced smaller radiative tendencies than those seen here (see their Fig. 5a), suggesting that vertical resolution accounts for most of the difference in radiative heating rates.}.
Time-mean radiative warming spread from 30- to 160-km radius within the cirrus canopy.
The existence of radiative cooling overlying radiative warming in this region led to radiatively-forced destabilization at and below the tropopause, as was observed in Fig. \ref{fig:stab-48-72}d.
Below the warming layer existed a region of forcing for stabilization, while a much stronger region of forcing for destabilization existed above the cooling maximum in the lower stratosphere.

@@

Although vertical turbulence always acts to eliminate vertical gradients of $\theta$, this adjustment toward a neutral state only occurs where the mixing takes place.
If turbulence occurs in a stably-stratified layer, it will act to decrease $\theta$ at the top of the layer and increase it below.
Just above and just below the mixed layer, however, the $\theta$ profile remains undisturbed.
Consequently, although turbulent mixing acts to decrease $\partial \theta/\partial z$ in the layer in which it is occurring, it actually increases $\partial \theta/\partial z$ just below and just above the layer.
These vertical gradients of turbulent mixing are quite important, particularly on the flanks of the upper-tropospheric outflow jet.

Diurnal variability of of static stability in the upper troposphere is an interesting area of future research.

Bu et al. did not see a lower-stratospheric inflow layer, likely due to their 625-m grid spacing. If we want to resolve properly the radiative cooling at the top of the cirrus canopy and the lower-stratospheric inflow layer, we must have sufficient vertical resolution in the tropopause layer. Our 100-m grid spacing here may not even be sufficient, given how shallow the layer is!
Did my no-CRF simulation show a lower-stratospheric inflow layer?? If not, that's strong evidence supporting the role of radiation in producing the desceding inflow layer above the tropopause.

@@

 \subsection{Sensitivity to the vertical mixing length}
The intensity of parameterized turbulence is highly dependent on a prescribed length scale.
Since there is no theoretical or observational guidance for the selection of this mixing length, the value used in the control run (100 m) is based on the sensitivity experiments of \cite{Bryan2012}.
Since the vertical eddy viscosity varies linearly with the vertical mixing length, prescribing a smaller mixing length produces smaller $\theta$ tendencies due to turbulence.
Even with a mixing length on the low end of those tested by \cite{Bryan2012}, however, turbulence still plays a role in the tropopause-layer $N^2$ evolution.
FIG shows the $N^2$ evolution in a simulation identical to the control run, except with a vertical mixing length of 50 m rather than 100 m.
DESCRIPTION OF THE FIGURE

@@

%FIGURE 5%
\begin{figure*}[ht]
\centerline{\includegraphics[width=27pc]{figures/h000-h024-budgetterms.png}}
\caption{(a) Total change in $N^2$ over the 0-24-hour period (10\textsuperscript{-4} s\textsuperscript{-2} (24 h)\textsuperscript{-1}) and the contributions to that change from (b) the sum of horizontal and vertical advection, (c) vertical turbulence, (d) longwave and shortwave radiation, (e) the sum of horizontal advection, vertical advection, and vertical turublence, and (f) the sum of horizontal advection, vertical advection, vertical turbulence, and longwave and shortwave radiation.}
\label{fig:stab-00-24}
\end{figure*}

%FIGURE 6%
\begin{figure*}[ht]
\centerline{\includegraphics[width=27pc]{figures/h024-h048-budgetterms.png}}
\caption{As in Fig.~\ref{fig:stab-00-24}, but for the 24-48-hour period.}
\label{fig:stab-24-48}
\end{figure*}

%FIGURE 7%
\begin{figure*}[ht]
\centerline{\includegraphics[width=27pc]{figures/h048-h072-budgetterms.png}}
\caption{As in Fig.~\ref{fig:stab-00-24}, but for the 48-72-hour period.}
\label{fig:stab-48-72}
\end{figure*}

@@

%FIGURE 11%
\begin{figure*}[ht]
\centerline{\includegraphics[width=27pc]{figures/qi+radten.png}}
\caption{Ice mixing ratio (g kg\textsuperscript{-1}) and cold-point tropopause height (orange lines) averaged over (a) 0-24 hours, (c) 24-48 hours, and (e) 48-72 hours. Radiative heating rate (K h\textsuperscript{-1}) and cold-point tropopause height (orange lines) averaged over (b) 0-24 hours, (d) 24-48 hours, and (f) 48-72 hours.} 
\label{fig:qi+radten}
\end{figure*}

>>>>

\abstract{Large changes in tropopause-layer static stability are observed during the rapid intensification (RI) of an idealized, axisymmetric tropical cyclone (TC).
Over the eye, static stability near the tropopause decreases and the cold-point tropopause height rises by up to 4 km at the storm center.
Outside of the eye, static stability increases considerably just above the cold-point tropopause, and the tropopause remains near its initial level.\\
\indent A budget analysis reveals that advection contributes to the static stability tendencies at all times throughout the upper troposphere and lower stratosphere.
Advection is particularly important within the eye, where it acts to destabilize the layer near and above the cold-point tropopause.
Outside of the eye, a radial-vertical circulation develops during RI, with strong outflow below the tropopause and weak inflow above.
Vertical wind shear above and below the upper-tropospheric outflow maximum induces turbulence, which provides forcing for both destabilization and stabilization in the tropopause layer.
Meanwhile, as organized convection reaches the tropopause, radiative heating tendencies at the top of the cirrus canopy generally act to destabilize the upper troposphere and stabilize the lower stratosphere.
Turbulent mixing and radiative heating combine to play an important role in the development of the strong stable layer immediately above the cold-point tropopause during RI.
The results suggest that turbulence and radiation, alongside advection, play fundamental roles in the upper-level static stability evolution of TCs.}
%The conclusions are robust to changes in the initial thermodynamic profile and prescribed vertical mixing length within a reasonable range of values.}

@@

Idealized simulations of a TC analyzed by \cite{OhnoSatoh2015} suggested that the development of an upper-level warm core near the TC storm center acted to decrease the static stability near the tropopause (compare their Figs. 9,10).
Although the mechanisms that drive this static stability evolution have not been examined explicitly, \cite{SternZhang2013} described the development of the TC warm core using a potential temperature ($\theta$) budget analysis.
They found that radial and vertical advection both played important roles in warm core development throughout RI, and subgrid-scale diffusion became particularly important during the later stage of RI.
To our knowledge, the only paper that has examined explicitly the static stability evolution in a modeled TC is \cite{Kepertetal2016}, but their analysis was limited to the boundary layer.
The analysis herein is based upon that of \cite{SternZhang2013}, except using a static stability budget similar to that of \cite{Kepertetal2016}, with a focus on the upper troposphere and lower stratosphere.

@@

In the tropopause layer, some of the budget terms are small enough to be ignored.
To determine which of the budget terms are most important, a time series of the contribution of each of the budget terms in Eq. \ref{eq:dthetadt} to the tropopause-layer static stability tendency is plotted in Fig. \ref{fig:avgbudterms}.
For this figure, each of the budget terms is computed using the method described in Section 3, except with 1-hour averaging intervals instead of 24-hour intervals.
The absolute values of these tendencies are then averaged over a radius-height domain surrounding the tropopause and plotted as a time series\footnote{It will be seeen in subsequent figures that each of the terms contributes both positively and negatively to the $N^2$ tendency within the analysis domain. 
Thus, taking an average over the domain tends to wash out the positive and negative contributions.
To circumvent this problem, the absolute value of each of the terms is averaged.}. 
Advection (Fig.~\ref{fig:avgbudterms}, red line) plays an important role in the mean tropopause-layer static stability tendency at all times, and vertical turbulence (Fig.~\ref{fig:avgbudterms}, blue line) and radiation (Fig.~\ref{fig:avgbudterms}, dark green line) also contribute significantly. %both become important after 48 hours.
Although the contribution from horizontal turbulence (Fig.~\ref{fig:avgbudterms}, purple line) becomes more important after 48 hours, it is confined to a very small region immediately surrounding the eyewall tangential velocity maximum (not shown), and is negligible throughout the rest of the tropopause layer.
The remaining two processes - microphysics and dissipative heating (Fig.~\ref{fig:avgbudterms}, orange and light green lines, respectively) - lie atop one another near zero.
Although the latent heating term can be quite large in the eyewall region, it is negligible everywhere outside of the eyewall, as are the effects of dissipative heating.
These time series indicate that, at all times, three budget terms dominate the tropopause-layer static stability tendency: advection, vertical turbulence, and radiation.
Variations in the magnitude and spatial structure of these terms drive the static stability changes depicted in Fig.~\ref{fig:mod+bud+res}; subsequent sections will focus on these variations and what causes them.

@@

\paragraph{0-24 hours}
The weakening of the lower-stratospheric static stability maximum during the initial spin-up period is reflected in the total $N^2$ budget change over this time (Fig.~\ref{fig:stab-00-24}a).
The 17-18-km layer was characterized by decreasing $N^2$ (purple shading), maximizing at the storm center.
The layer immediately below the tropopause, meanwhile, saw strengthening $N^2$ during this time period.
Although these tendencies extended out to the 200-km radius, they were particularly pronounced at innermost radii.
A comparison of the contributions of advection (Fig.~\ref{fig:stab-00-24}b), vertical turbulence (Fig.~\ref{fig:stab-00-24}c), and radiation (Fig.~\ref{fig:stab-00-24}d) reveals that advection was primarily responsible for the change in static stability during this period.
Although vertical turbulence acted in opposition to advection (i.e. it acted to stabilize regions that advection acted to destabilize), the magnitude of the advective tendencies was larger, particularly at the innermost radii.
The sum of advection and vertical turbulence (Fig.~\ref{fig:stab-00-24}e) almost exactly replicated the static stability tendencies above 17 km.
Radiative tendencies. meanwhile, (Fig.~\ref{fig:stab-00-24}d) acted to destabilize the layer below about 16 km and stabilize the layer between 16 and 17 km.
The sum of advection, vertical turbulence, and radiation (Fig.~\ref{fig:stab-00-24}f) reproduces the total change in $N^2$ almost exactly.

@@

The sum of advection, vertical turbulence, and radiation (Fig. \ref{fig:stab-24-48}f) once again closely follows the observed $N^2$ variability, except in portions of the eyewall where the neglect of latent heating and horizontal turbulence introduces some differences.

@@

Advection played an important role in the tropopause-layer $N^2$ evolution at all stages of intensification, but for brevity, this section will focus only on the RI (24-48-hour) period.
To investigate the advective processes more closely, the individual contributions of horizontal and vertical advection during the RI period are shown in Fig. \ref{fig:adv-24-48}, along with the corresponding time-mean radial and vertical velocities and $\theta$.
The $N^2$ tendencies due to the two advective components (Fig. \ref{fig:adv-24-48}a,b) exhibit strong cancellation, consistent with flow that is nearly isentropic.
There are, however, many regions in which flow crosses $\theta$ surfaces; this flow accounts for all non-zero $N^2$ tendencies due to advection previously seen in Fig. \ref{fig:stab-24-48}b.

%The eye was characterized by stabilization below 16 km and destabilization in the 16-17-km and 18-19-km layers.
%Comparing Fig. \ref{fig:stab-24-48}b to \ref{fig:adv-24-48}a,b reveals that the stabilization below 16 km in the eye was forced by vertical advection, as was the region of destabilization in the 18-19-km layer.
%Meanwhile, the destabilization that occurred in the 16-17-km layer was forced by a combination and radial and vertial advection.

%Outside of the eye, the layer below about 15.5 km was dominated by stabilization.
%Outside of the 50-km radius, horizontal advection provided the forcing for this stabilization, while inside of the 50-km radius, vertical advection was responsible.
%Above this layer existed a layer of destabilization near the 16-km level, which was forced by horizontal advection.

@@

The first two terms on the right-hand side of Eq. \ref{eq:advtend} represent advection of static stability by the radial and vertical wind, respectively.
These terms cannot create a maximum or a minimum; they only act to rearrange the static stability field.
%Since the goal of this analysis is to understand the processes responsible for the creation of maxima and minima in $N^2$, these two terms will not be examined.
The third and fourth terms represent, respectively, the tilting of isentropes in the presence of vertical wind shear, and the spreading or compaction of isentropes through divergence of the vertical wind.
Since these terms involve gradients of velocities, they can create or eliminate local maxima or minima in static stability.% This section will focus on these two terms.

@@

%   \paragraph{RI period: 24-48 hours}
During the RI period, strong radial and vertical circulations developed near the tropopause, which forced high-magnitude $N^2$ tendencies due to advection (Fig. \ref{fig:adv-24-48}a,b).
A layer of strong outflow developed at and below the tropopause during this period, with the outflow maximum (dashed cyan line) curving from the 14-km level at the 50-km radius to just below the 16-km level outside of the 80-km radius (Fig. \ref{fig:adv-24-48}c).
The cyan line, by definition, represents the level at which the vertical gradient of radial velocity switched signs, with $\partial u/\partial z>0$ below the line and $\partial u/\partial z<0$ above.
Notably, the $N^2$ tendency due to horizontal advection (Fig. \ref{fig:adv-24-48}a) also tended to switch signs at this line, with stabilization below the outflow maximum and destabilization above.
%This suggests that vertical wind shear above and below the outflow maximum played an important role in the $N^2$ tendency during this time.
Examination of the third term on the right-hand side of Eq. \ref{eq:advtend} suggests that tilting of isentropes within these shear layers contributed to the strong $N^2$ tendencies that flanked the outflow maximum.
Outside of the eye and eyewall, isentropes generally sloped upward with radius, which means that $\theta$ decreased outward ($\partial \theta/\partial r<0$).
Thus, wherever $\partial u/\partial z>0$, the tilting term must force an increase in $N^2$, and wherever $\partial u/\partial z<0$, the tilting term must force a decrease in $N^2$, which is precisely the structure seen in Fig. \ref{fig:adv-24-48}a.

Advection of $N^2$ by the radial wind (first term on the right-hand side of Eq. \ref{eq:advtend}) also acted within the outflow jet.
For example, horizontal advection provided forcing for destabilization at the 16-km level almost everywhere inside of the 140-km radius.
Outside of this radius near 16 km, however, existed a region of forcing for stabilization.
This switch in signs was a consequence of a reversal of the radial graident of mean $N^2$ near the 140-km radius (Fig. \ref{fig:n2-24hr-avgs}b).
Inside of that radius, $(\partial/\partial r)(\partial \theta/\partial z) > 0$ and $u > 0$, which corresponds to forcing for destabilization in Eq. \ref{eq:advtend}; outside of that radius $(\partial/\partial r)(\partial \theta/\partial z) < 0$ and $u > 0$, which corresponds to forcing for stabilization.

@@

Meanwhile in the lower stratosphere, a thin layer of 2-4 m s\textsuperscript{-1} inflow developed a few hundred meters above the tropopause, similar to that which was observed in Hurricane Patricia (2015; \citeauthor{DuranMolinari2018} \citeyear{DuranMolinari2018}) and in previous modeling studies (e.g. \citeauthor{OhnoSatoh2015} \citeyear{OhnoSatoh2015}; \citeauthor{Kieuetal} \citeyear{Kieuetal}).
Since the isentropes in this layer sloped slightly upward with radius (i.e. $\partial \theta/\partial r < 0$), this inflow acted to import lower $\theta$ air from outer radii to inner radii.
%There also existed strong radial gradients of mean $N^2$ in this layer during this time (Fig. \ref{fig:n2-24hr-avgs}b), so the inflow layer must have acted to rearrange $N^2$.
%This rearrangement could not have been responsible for the overall increase in $N^2$ during this period, however.
%If horizontal advection contributed to the strengthening stable layer just above the tropopause, it must have been through the tilting term.
%The inflow maximum near the 18-km level represents the level at which the sign of $\partial u/\partial z$ reverses, with $\partial u/\partial z < 0$ below and $\partial u/\partial z > 0$ above.
%The third term on the right-hand side of Eq. \ref{eq:advtend} shows that the tilting of isentropes should act to destabilize the layer below the inflow maximum and stabilize the layer above, which is precisely what was observed outside of the 80-km radius in Fig. \ref{fig:adv-24-48}a.
Decreasing $\theta$ near the inflow maximum acted to destabilize the layer below the inflow maximum and stabilize the layer above (Fig. \ref{fig:adv-24-48}a).

Vertical advection also played an important role in the tropopause-layer static stability evolution.
Within the eye, subsidence dominated below 17 km, while mean ascent existed near the storm center above 17 km\footnote{Close examination of the model output revealed that this mean ascent was forced by the interaction of inward-propagating waves with the boundary at r=0. When a wave reached r=0, a dipole of vertical velocity resulted, with ascent above and subsidence below. For reasons that remain unclear, the regions of ascent were more persistent than the regions of subsidence, which resulted in mean ascent near r=0 in the lower stratosphere.}.
Although the magnitude of the subsidence was larger at lower altitudes, $\partial \theta/\partial z$ was smaller there.
Because $\partial \theta/\partial z$ was smaller, the subsidence at lower levels could not accomplish as much warming as the subsidence at higher levels in the eye, consistent with the results of \cite{SternZhang2013}.

Outside of the eye, ascent dominated the troposphere, while a 1.5-km-deep layer of descent existed immediately above the tropopause.
This region of descent

@@

After the TC's RI period completed (48-72 hours; Fig. \ref{fig:qi+radten}f), strong radiative cooling remained near the tropopause at inner radii, sloping downward with the top of the cirrus canopy to below the tropopause at outer radii.
Cooling rates exceeded 1 K h\textsuperscript{-1} just above the tropopause between the 30- and 70-km radii.
These cooling rates exceeded those observed by \cite{buetal}, a discrepancy that is a consequence of their larger vertical grid spacing (625 m) compared to that used here (250 m), along with a contribution from differing radiation schemes\footnote{\cite{buetal} employed the NASA-Goddard radiation scheme for their CM1 simulations, whereas RRTMG is used in the present paper. A simulation using NASA-Goddard radiation and 625-m vertical grid spacing produced maximum radiative cooling rates of 0.3 K h\textsuperscript{-1}, which agrees with the rates in \cite{buetal}. Another simulation using 625-m vertical grid spacing and RRTMG radiation produced cooling rates of up to 0.6 K h\textsuperscript{-1}.}.
Time-mean radiative warming spread from 30- to 160-km radius within the cirrus canopy.
The existence of radiative cooling overlying radiative warming in this region led to radiatively-forced destabilization at and below the tropopause, as was observed in Fig. \ref{fig:stab-48-72}d.
Below the warming layer existed a region of forcing for stabilization, while a much stronger region of forcing for destabilization existed above the cooling maximum in the lower stratosphere.

@@

%An interesting result of the budget analysis was that turbulent mixing acted to increase static stability in some regions of the storm.
Although vertical turbulence always acts to eliminate vertical gradients of $\theta$, this adjustment toward a neutral state only occurs where the mixing takes place.
If turbulence occurs in a stably-stratified layer, it will act to decrease $\theta$ at the top of the layer and increase it below.
Just above and just below the mixed layer, however, the $\theta$ profile remains undisturbed.
Consequently, although turbulent mixing acts to decrease $\partial \theta/\partial z$ in the layer in which it is occurring, it actually increases $\partial \theta/\partial z$ just below and just above the layer.
These vertical gradients of turbulent mixing are quite important, particularly on the flanks of the upper-tropospheric outflow jet.

Fig. \ref{fig:diff} reveals that two distinct maxima of vertial eddy diffusivity developed in the tropopause layer as the storm intensified.
Comparison of these turbulent regions to the $N^2$ tendencies in Figs. \ref{fig:stab-24-48}c and \ref{fig:stab-48-72}c reveals that the layers in which vertical eddy diffusivity maximized corresponded to layers of destabiliztion due to vertical turbulence.
Just outside of these layers, however, vertical turbulence acted to increase $N^2$.
The large vertical gradient of vertical eddy diffusivity near the tropopause played an important role in developing the lower-stratospheric stable layer during RI.
This supports the hypothesized role of turbulence in setting the outflow-layer $\theta$ stratification in \cite{RotunnoEmanuel}.

  \section{Conclusions}

The simulated $N^2$ evolution shown herein closely matched that observed during the RI of Hurricane Patricia (2015).
Three processes dominated the $N^2$ variability in the upper troposphere and lower stratosphere: advection, radiation, and vertical turbulence.
Radiation and vertical turbulence played particularly important roles in developing the strong $N^2$ maximum just above the cold-point tropopause during RI.
Since these two processes are parameterized, and radiation closely depends on yet another parameterized process (microphysics), the tropopause-layer $N^2$ variability could be quite sensitive to the assumptions inherent to parameterizations used.
A number of sensitivity experiments revealed that the details of the $N^2$ evolution were sensitive to the selection of the microphysics scheme, the vertical turbulence length scale, and the vertical grid spacing.
All of the simulations, however, did develop an elevated tropopause over the eye that corresponded to decreasing static stability and, outside of the eye, an $N^2$ maximum immediately above the tropopause.

Comparison to previous work revealed that small vertical grid spacing is necessary to resolve the shallow layer of radiative cooling at the top of the cirrus canopy.
Increasing the vertical grid spacing decreased the magnitude of the radiative cooling and produced a weaker, but deeper lower-stratospheric stable layer.

In this paper, all of the variables were averaged over a full diurnal cycle to eliminate the effects of diurnal variability and isolate the overall storm evolution.
Diurnal variations in static stability near the tropopause are potentially of interest with respect to the tropical cyclone diurnal cycle, however, and will be the subject of future work.

@@

As an example, 24-hour averages of $N^2$ are plotted in Fig. \ref{fig:n2-1000km} for a simulation that was identical to that used in this paper, except the initial sounding was determined by averaging every grid point within 1000 km of TC Patricia's storm center at 18 UTC 21 October 2015 instead of averaging only within the 100-km radius.
Although the lower-stratospheric stable layer developed more slowly and was weaker than that shown in Fig. \ref{fig:n2-24hr-avgs}, the overall evolution was quite similar and the same budget terms dominated the static stability evolution (not shown).

 \subsection{Sensitivity to the vertical mixing length}
The rate of turbulent mixing in the Smagorinsky scheme used herein is highly dependent on a prescribed length scale.
The vertical mixing length used in this paper (100 m) was based on the sensitivity experiments of \cite{Bryan2012}.
Prescribing a smaller mixing length produces smaller $\theta$ tendencies due to turbulence, but even with a mixing length on the low end of those tested by \cite{Bryan2012}, turbulence still played an important role in the tropopause-layer $N^2$ evolution.
Fig. \ref{fig:turb-50m} shows the 24-hour-averaged contributions of turbulent mixing to the $N^2$ evolution from a simulation identical to that used in this paper, except with a vertical mixing length of 50 m.
At all times, vertical turbulence still played an important role in the tropopause-layer static stability evolution, particularly during the latter stages of RI (48-72 hours).

@@

%FIGURE 5%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/h000-h024-budgetterms.png}}
\end{figure*}
\begin{figure}
\caption{(a) Total change in $N^2$ over the 0-24-hour period (10\textsuperscript{-4} s\textsuperscript{-2} (24 h)\textsuperscript{-1}) and the contributions to that change from (b) the sum of horizontal and vertical advection, (c) vertical turbulence, (d) longwave and shortwave radiation, (e) the sum of horizontal advection, vertical advection, and vertical turublence, and (f) the sum of horizontal advection, vertical advection, vertical turbulence, and longwave and shortwave radiation.}
\label{fig:stab-00-24}
\end{figure}

%FIGURE 6%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/h024-h048-budgetterms.png}}
\caption{As in Fig.~\ref{fig:stab-00-24}, but for the 24-48-hour period.}
\label{fig:stab-24-48}
\end{figure*}

%FIGURE 7%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/h048-h072-budgetterms.png}}
\caption{As in Fig.~\ref{fig:stab-00-24}, but for the 48-72-hour period.}
\label{fig:stab-48-72}
\end{figure*}

@@

%FIGURE 9%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/qi+radten.png}}
\end{figure*}
\begin{figure}
\caption{Ice mixing ratio (g kg\textsuperscript{-1}) and cold-point tropopause height (orange lines) averaged over (a) 0-24 hours, (c) 24-48 hours, and (e) 48-72 hours. Radiative heating rate (K h\textsuperscript{-1}) and cold-point tropopause height (orange lines) averaged over (b) 0-24 hours, (d) 24-48 hours, and (f) 48-72 hours.} 
\label{fig:qi+radten}
\end{figure}
